---
title: "exercise-11"
format: html
editor: visual
---

# Preliminaries

``` {r}
library(tidyverse)
library(skimr)
d <- read_tsv("https://raw.githubusercontent.com/difiore/ada-datasets/main/Mammal_lifehistories_v2.txt")
d[d == -999] <- NA # removes "-999" values and properly replaces them with "NA"
skim(d)
```

``` {r}
head(d)
```

The values order, family, Genus, and species are categorical, while mass(g), gestation(mo), newborn(g), weaning(mo), wean mass(g), AFR(mo), max.life(mo), litter size, litters/year, and refs are numeric.

# Challenge

## Step 1

``` {r}
d[d == -999] <- NA # I already did this part but I'll just put it here again
```

## Step 2

``` {r}
d <- d %>%
  dplyr::select(-c("litter size", "refs"))
d <- d %>%
  mutate(`maxlife(mo)` = `max. life(mo)`, `weanmass(g)` = `wean mass(g)`) %>%
  dplyr::select(!c("max. life(mo)", "wean mass(g)"))
head(d)
```

## Step 3

``` {r}
d <- d %>%
  mutate(across(where(is.numeric) & !c("maxlife(mo)", "weanmass(g)"), log))
head(d)
```

## Step 4

``` {r}
m1 <- lm(data = d, `mass(g)` ~ `gestation(mo)`, na.action = na.exclude)
relgest <- resid(m1)
m2 <- lm(data = d, `mass(g)` ~ `weaning(mo)`, na.action = na.exclude)
relwean <- resid(m2)
m3 <- lm(data = d, `mass(g)` ~ `AFR(mo)`, na.action = na.exclude)
relafr <- resid(m3)
m4 <- lm(data = d, `mass(g)` ~ `maxlife(mo)`, na.action = na.exclude)
rellife <- resid(m4)
m5 <- lm(data = d, `mass(g)` ~ `newborn(g)`, na.action = na.exclude)
relnewbornmass <- resid(m5)
m6 <- lm(data = d, `mass(g)` ~ `weanmass(g)`, na.action = na.exclude)
relweaningmass <- resid(m6)
d <- d %>%
  mutate(relGest = relgest, relWean = relwean, relAFR = relafr, relLife = rellife, relNewbornMass = relnewbornmass, relWeaningMass = relweaningmass)
head(d)
```

## Step 5

``` {r}
ggplot(d, aes(order, relLife)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

Artiodactyla and Perissodactyla appear to have the highest residual lifespan.

``` {r}
ggplot(d, aes(order, relNewbornMass)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

Carnivora appears to have the highest residual newborn mass; many more, like Lagomorpha, Perissodactyla, Proboscidea, and Tubulidentata appear to be high in the running as well.

``` {r}
ggplot(d, aes(order, relWeaningMass)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

Proboscidea, Sirenia, and Perissodactyla appear to have the highest residual weaning mass.

## Step 6

``` {r}
d_maxlife <- d %>%
  drop_na(`maxlife(mo)`, `gestation(mo)`, `newborn(g)`, `weaning(mo)`, `weanmass(g)`, `litters/year`, `mass(g)`)
library(MASS)
m1 <- lm(data = d_maxlife, `maxlife(mo)` ~ 1)
s1 <- stepAIC(m1, scope = .~. + `gestation(mo)` + `newborn(g)` + `weaning(mo)` + `weanmass(g)` + `litters/year` + `mass(g)`, direction = "both")
```

For maxlife(mo), the best overall model is the one containing nearly all predictors (mass(g), weanmass(g), weaning(mo), litters/year, and gestation(mo)). It looks like 6 candidate models (mass(g) + newborn(g), mass(g) + weanmass(g) + newborn(g), mass(g) + weanmass(g) + weaning(mo) + newborn(g), mass(g) + weanmass(g) + weaning(mo) + gestation(mo), mass(g) + weanmass(g) + weaning(mo) + litters/year + newborn(g), and mass(g) + weanmass(g) + weaning(mo) + litters/year + gestation(mo) + newborn(g)) have a delta AIC of 4 or less.

``` {r}
d_maxafr <- d %>%
  drop_na(`AFR(mo)`, `gestation(mo)`, `newborn(g)`, `weaning(mo)`, `weanmass(g)`, `litters/year`, `mass(g)`)
m2 <- lm(data = d_maxafr, `AFR(mo)` ~ 1)
s2 <- stepAIC(m2, scope = .~. + `gestation(mo)` + `newborn(g)` + `weaning(mo)` + `weanmass(g)` + `litters/year` + `mass(g)`, direction = "both")
```

For maxAFR(mo), the best overall model is the one containing 4 predictors (litters/year, gestation(mo), mass(g), and weaning(mo)). It looks like 6 candidate models (mass(g) + newborn(g), mass(g) + weanmass(g) + newborn(g), mass(g) + weanmass(g) + weaning(mo) + newborn(g), mass(g) + weanmass(g) + weaning(mo) + gestation(mo), mass(g) + weanmass(g) + weaning(mo) + litters/year + newborn(g), and mass(g) + weanmass(g) + weaning(mo) + litters/year + gestation(mo) + newborn(g)) have a delta AIC of 4 or less. However, I did notice that stepAIC added mass(g) to  AFR(mo) ~ litters/year + gestation(mo) even though the delta AIC for that candidate model was -331.13 - -334.58 = 3.45 < 4, so maybe AFR(mo) ~ litters/year + gestation(mo) should be the best overall model instead.

mass(g), weaning(mo), litters/year, and gestation(mo) appear in both top models (however, if we accept AFR(mo) ~ litters/year + gestation(mo) as the *real* top model for maxAFR(mo), then it's just litters/year and gestation(mo)).

``` {r}
d_maxlife <- d %>%
  drop_na(`maxlife(mo)`, `gestation(mo)`, `newborn(g)`, `weaning(mo)`, `weanmass(g)`, `litters/year`, `mass(g)`)
maxlife0 <- lm(data = d_maxlife, `maxlife(mo)` ~ 1)
summary(maxlife0)
```

``` {r}
add1(maxlife0, scope = .~. + `gestation(mo)` + `newborn(g)` + `weaning(mo)` + `weanmass(g)` + `litters/year` + `mass(g)`, test = "F")
```

``` {r}
maxlife1 <- update(maxlife0, formula = .~. + `mass(g)`)
summary(maxlife1)
```

``` {r}
add1(maxlife1, scope = .~. + `gestation(mo)` + `newborn(g)` + `weaning(mo)` + `weanmass(g)` + `litters/year`, test = "F")
```

``` {r}
maxlife2 <- update(maxlife1, formula = .~. + `weanmass(g)`)
summary(maxlife2)
```

``` {r}
add1(maxlife2, scope = .~. + `gestation(mo)` + `newborn(g)` + `weaning(mo)` + `litters/year`, test = "F")
```

``` {r}
maxlife3 <- update(maxlife2, formula = .~. + `weaning(mo)`)
summary(maxlife3)
```

``` {r}
add1(maxlife3, scope = .~. + `gestation(mo)` + `newborn(g)` + `litters/year`, test = "F")
```

``` {r}
maxlife4 <- update(maxlife3, formula = .~. + `litters/year`)
summary(maxlife4)
```

``` {r}
add1(maxlife4, scope = .~. + `gestation(mo)` + `newborn(g)`, test = "F")
```

``` {r}
maxlife5 <- update(maxlife4, formula = .~. + `gestation(mo)`)
summary(maxlife5)
```

``` {r}
add1(maxlife5, scope = .~. + `newborn(g)`, test = "F")
```
``` {r}
library(MuMIn)
maxlife <- lm(data = d_maxlife, `maxlife(mo)` ~ `mass(g)` + `weanmass(g)` + `weaning(mo)` + `litters/year` + 
    `gestation(mo)`, na.action = "na.fail")
maxlife_mods <- dredge(maxlife, beta = "none", m.lim = c(0, 4))
maxlife_mods_list <- get.models(maxlife_mods, subset = TRUE)
coef(maxlife_mods)
```

``` {r}
maxlife_mods_avg <- summary(model.avg(maxlife_mods, subset = delta < 4, fit = TRUE))
maxlife_mods_avg
```

``` {r}
confint(maxlife_mods_avg)
```

``` {r}
plot(maxlife_mods_avg, full = TRUE, intercept = FALSE)
```

``` {r}
plot(maxlife_mods_avg, full = FALSE, intercept = FALSE)
```

``` {r}
maxafr0 <- lm(data = d_maxafr, `AFR(mo)` ~ 1)
summary(maxafr0)
```
``` {r}
add1(maxafr0, scope = .~. + `gestation(mo)` + `newborn(g)` + `weaning(mo)` + `weanmass(g)` + `litters/year` + `mass(g)`, test = "F")
```

``` {r}
maxafr1 <- update(maxafr0, formula = .~. + `litters/year`)
summary(maxafr1)
```

``` {r}
add1(maxafr1, scope = .~. + `gestation(mo)` + `newborn(g)` + `weaning(mo)` + `weanmass(g)` + `mass(g)`, test = "F")
```

``` {r}
maxafr2 <- update(maxafr1, formula = .~. + `gestation(mo)`)
summary(maxafr2)
```

``` {r}
add1(maxafr2, scope = .~. + `newborn(g)` + `weaning(mo)` + `weanmass(g)` + `mass(g)`, test = "F")
```

``` {r}
maxafr3 <- update(maxafr2, formula = .~. + `mass(g)`)
summary(maxafr3)
```

``` {r}
add1(maxafr3, scope = .~. + `newborn(g)` + `weaning(mo)` + `weanmass(g)`, test = "F")
```

``` {r}
maxafr4 <- update(maxafr3, formula = .~. + `weaning(mo)`)
summary(maxafr4)
```

``` {r}
add1(maxafr4, scope = .~. + `newborn(g)` + `weanmass(g)`, test = "F")
```

``` {r}
maxafr <- lm(data = d_maxafr, `AFR(mo)` ~ `litters/year` + `gestation(mo)` + `mass(g)` + `weaning(mo)`, na.action = "na.fail")
maxafr_mods <- dredge(maxafr, beta = "none", m.lim = c(0, 4))
maxafr_mods_list <- get.models(maxafr_mods, subset = TRUE)
coef(maxafr_mods)
```

``` {r}
maxafr_mods_avg <- summary(model.avg(maxafr_mods, subset = delta < 5, fit = TRUE)) # the lowest delta in this model is 4.88 so I had to increase the acceptable subset
confint(maxafr_mods_avg)
```

``` {r}
plot(maxafr_mods_avg, full = TRUE, intercept = FALSE)
```

``` {r}
plot(maxafr_mods_avg, full = FALSE, intercept = FALSE)
```